---
title: "Implementing cookie consent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Implementing cookie consent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Cookie consent is a mandatory requirement for any dashboard when using cookies. 
Typically, all DfE dashboards will be using cookies as part of tracking with
Google Analytics and so will require cookie consent.

> Note that to activate Google Analytics on your data dashboard, you will need 
to contact the Statistics Development team at 
[statistics.development@education.gov.uk](mailto:statistics.development@education.gov.uk) and request a Google Analytics key.

dfeshiny provides a standard GDS compliant banner, plus the underlying code, to
allow end users to give or withhold their consent for cookies to be used. If 
they withhold their consent, then tracking with Google Analytics will be halted.

Implementing cookie control on your dashboard is handled by several functions 
in dfeshiny, with the process outlined below.

The cookie control requires some underlying javascript to remove and create 
cookies as well as switch Google Analytics on and off. This javascript needs 
including in your dashboard's www/ folder, which can be achieved by running the 
following command from the R console:

```{r eval=FALSE}
dfeshiny::init_cookies()
```

Before adding the cookie consent banner, we recommend adding your Google 
Analytics key as a variable to your dashboard's [global.R](https://github.com/dfe-analytical-services/dfeshiny/blob/cookie-module/tests/test_dashboard/global.R) 
file (replacing `ABCDE12345` below with the key provided by the Statistics 
Development team):

```
google_analytics_key <- 'ABCDE12345'
```

You will then need to add the necessary code to your ui.R file to a) run the 
javascript, b) create the consent banner and c) add a cookies information panel:


```{r eval=FALSE}
ui <- function(input, output, session) {
  fluidPage(
    shinyjs::useShinyjs(),
    dfe_cookie_script(),
    cookie_banner_ui("Your dashboard name"),
    ...
    cookies_panel()
  )
}
```

You should then add the `cookie_banner_server` function to your server.R script:

```{r eval=FALSE}
cookie_banner_server(
   "cookies",
   input_cookies = reactive(input$cookies),
   input_clear = reactive(input$cookie_consent_clear),
   parent_session = session,
   google_analytics_key = google_analytics_key
)
```

In the above you should keep all of the flags as they are shown. Note that this 
assumes you have added `google_analytics_key` as a declared variable in your 
global.R file. If you've not done this, you'll need to replace 
`google_analytics_key = google_analytics_key` to 
`google_analytics_key = "ABCDE12345"` (replacing "ABCDE12345" with your own 
key).

With the above included, when the user's preference will itself be stored in a
cookie (`dfe_analytics`), which by default expires after 365 days. The user can
then also find information about the cookies stored by your dashboard and change 
their consent preference on a specific cookie consent panel on your dashboard 
that's created by the `cookies_panel()` function.
