---
title: "Implementing cookie consent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Implementing cookie consent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Cookie consent is a mandatory requirement for any dashboard when using cookies. 
Typically, all DfE dashboards will be using cookies as part of tracking with
Google Analytics and so will require cookie consent.

dfeshiny provides a standard GDS compliant banner, plus the underlying code, to
allow end users to give or withhold their consent for cookies to be used. If 
they withhold their consent, then tracking with Google Analytics will be halted.

Implementing cookie control on your dashboard is handled by several functions 
in dfeshiny, with the process outlined below.

The cookie control requires some underlying javascript to remove and create 
cookies as well as switch Google Analytics on and off. This javascript needs 
including in your dashboard's www/ folder, which can be achieved by running the 
following command from the R console:

```{r eval=FALSE}
dfeshiny::init_cookies()
```

You will then need to add the necessary code to your ui.R file to a) run the 
javascript, b) create the consent banner and c) add a cookies information panel:


```{r eval=FALSE}
ui <- function(input, output, session) {
  shinyjs::useShinyjs(),
  dfe_cookie_script()
  cookie_banner_ui("Your dashboard name"),
  shiny::navlistPanel(
    "",
    id = "navlistPanel",
    widths = c(2, 8),
    well = FALSE,
    dashboard_panel(),
    homepage_panel(),
    a11y_panel(),
    support_panel(
      team_email = "publication.team@education.gov.uk",
      repo_name = "https://github.com/dfe-analytical-services/your-dashboard-name",
      form_url = "https://forms.office.com"
    ),
    cookie_panel()
  )
}
```

You should then add the `cookie_banner_server` function to your server.R script:

```{r eval=FALSE}
cookie_banner_server(
   "cookies",
   input_cookies = reactive(input$cookies),
   input_clear = reactive(input$cookie_consent_clear),
   parent_session = session,
   google_analytics_key = "ABCDE12345"
)
```

In the above you should keep all of the flags as they are shown except for 
`google_analytics_key`, which you should update to your own 10 digit GA4 code.

