---
title: "Implementing cookies consent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Implementing cookies consent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Cookies are chunks of data that are stored by browsers to help websites store 
user preferences and other information. Most of our DfE data dashboards 
generally store two cookies, one to allow Google Analytics to differentiate 
between unique users and returning users and the second to store a user's 
preference on whether they allow the use of the Google Analytics (or any 
additional custom) cookies.

Assuming a dashboard is set up with at least the Google Analytics cookie, then 
gaining user consent is a mandatory requirement.

> Note that to activate Google Analytics on your data dashboard, you will need 
to contact the explore education statistics platforms team at 
[explore.statistics@education.gov.uk](mailto:explore.statistics@education.gov.uk) and request a Google Analytics key.

`dfeshiny` provides a standard Government Digital Service (GDS) compliant 
banner, a cookies information panel and the underlying code required to allow end 
users to give or withhold their consent for cookies to be used. If they withhold 
their consent, then tracking with Google Analytics will be halted (note that 
this will lead to your analytics underestimating your user base to some degree).

Implementing cookies control on your dashboard is handled by several functions 
in `dfeshiny`, with the process outlined below.

The cookies control requires some underlying javascript to remove and create 
cookies as well as switch Google Analytics on and off. This javascript needs 
including in your dashboard's www/ folder, which can be achieved by running the 
following command from the R console in your dashboard project / repo:

```{r eval=FALSE}
dfeshiny::init_cookies()
```

Before adding the cookies consent banner, we recommend adding your Google 
Analytics key as a variable to your dashboard's **[global.R](https://github.com/dfe-analytical-services/dfeshiny/blob/tests/test_dashboard/global.R)**
file (replacing `ABCDE12345` below with the key provided by the Explore Education Statistics Platforms team):

```{r eval=FALSE}
google_analytics_key <- "ABCDE12345"
```

You will then need to add the necessary code to your **ui.R** file to a) run the 
javascript (`dfe_cookies_script`), b) create the consent banner 
(`cookies_banner_ui`), editing the dashboard name to match your own and c) add a cookies information panel 
(`cookies_panel_ui`):


```{r eval=FALSE}
ui <- function(input, output, session) {
  shiny::fluidPage(
    shinyjs::useShinyjs(),
    dfeshiny::dfe_cookies_script(),
    dfeshiny::cookies_banner_ui(
      name = "Your dashboard name"
    ),
    shiny::navlistPanel(
      "",
      id = "navlistPanel",
      widths = c(2, 8),
      well = FALSE,
      shiny::tabPanel(
        value = "cookies_panel_ui",
        "Cookies",
        cookies_panel_ui(google_analytics_key = google_analytics_key)
      )
    )
  )
}
```

You should then add the `cookies_banner_server` and `cookies_panel_server` 
functions to your **server.R** script:

```{r eval=FALSE}
output$cookies_status <- dfeshiny::cookies_banner_server(
  input_cookies = shiny::reactive(input$cookies),
  parent_session = session,
  google_analytics_key = google_analytics_key
)

dfeshiny::cookies_panel_server(
  input_cookies = shiny::reactive(input$cookies),
  google_analytics_key = google_analytics_key
)
```

The ID for the UI panels is `navlistPanel`, and this is the default setting in the _server functions, if this ID is different, you will need to set cookie_nav_id in `cookies_panel_server()` to match your ID in the UI.

In the above you should keep all of the flags as they are shown. Note that this 
assumes you have added `google_analytics_key` as a declared variable in your 
global.R file. If you've not done this, you'll need to replace 
`google_analytics_key = google_analytics_key` to 
`google_analytics_key = "ABCDE12345"` (replacing "ABCDE12345" with your own 
key).

With the above included, the user's preference will itself be stored in a
cookie (`dfe_analytics`), which by default expires after 365 days. The user can
then also find information about the cookies stored by your dashboard and change 
their consent preference on a specific cookies consent panel on your dashboard 
that's created by the `cookies_panel()` function.
