---
title: "Using Air to format R code"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using Air to format R code}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

[Air](https://posit-dev.github.io/air/) is an R code formatter developed by
Posit. It automatically formats your R code to a consistent style, similar to
how `styler` works but faster and with deeper editor integration. Using Air
helps keep code readable and consistent across a team without manual effort.

Previously we used to use a custom `tidy_code()` function, and pre-commit
hooks to format the code, on bigger repositories this was painfully slow and
could add a minute or more to each commit. Using Air removes this and 
streamlines the process.

There are several ways to use Air, ranging from full local setup to relying on
a GitHub Actions workflow. 

We recommend setting up Air locally where possible, and using the format on 
save feature available in Positron, as this works incredibly well and will 
mean you never have to think about code formatting again.

We recommend every repo should have the GitHub workflow for two reasons:
1. All PRs (and therefore code changes) will be tested for formatting
2. Everyone can contribute without needing to go through any hoops of local setup

::: {.callout-warning}
Be aware, if you are adding Air to an existing repository you are likely to make
a lot of changes at once. It is best to do that on a clean PR, and when there
are no other open PRs that you are likely to have conflicts with.
:::

## Option 1: Positron (local)

[Positron](https://positron.posit.co/) is a next-generation IDE for data
science from Posit, with Air built in as the default R formatter. No extension
is needed.

If you are working in dfeshiny, format on save is already configured
for you via the `.vscode/settings.json` file at the root of the project.

To enable this in your own projects, open your user or workspace settings
(`Ctrl+Shift+P` > "Open User Settings (JSON)") and add:

```json
{
  "[r]": {
    "editor.defaultFormatter": "Posit.air-vscode",
    "editor.formatOnSave": true
  }
}
```

You'll then find that whenever you edit then save a file, the formatting is
automatically applied. It's blazing fast so be careful, if you blink you'll
miss it.

## Option 2: R Studio (local)

Air can be used in R Studio via the `usethis` package. Run the following once
from the R console in your project:

```{r eval=FALSE}
usethis::use_air()
```

This sets up Air for your project. Once configured, you can format the current
file using `Ctrl+Shift+A` (or `Cmd+Shift+A` on Mac) via the R Studio Addins
menu, or format your entire project by running:

```{r eval=FALSE}
air::format_project()
```

## Option 3: GitHub Actions workflow (every project)

For an additional safeguard, or for those who do not have Air set up locally,
we provide a reusable GitHub Actions workflow that checks R code formatting
on every pull request. Any formatting issues are surfaced as suggested changes
directly on the pull request, so they can be applied directly in GitHub
without needing Air installed locally.

To enable this in your repository, copy the following into a `.yaml` file in
your `.github/workflows/` folder:

```yaml
on:
  push:
    branches: [main, master]
  pull_request:

name: air

permissions:
  contents: read
  packages: read
  pull-requests: write

jobs:
  format-check:
    uses: dfe-analytical-services/dfeshiny/.github/workflows/air_reusable.yaml@main
```

Once added, the workflow will run on every pull request and push to main. If
any files are not formatted correctly, suggested fixes will appear as review
comments on the pull request. Accepting the suggestions will apply the
formatting automatically.
