---
title: "What to avoid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What to avoid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(dfeshiny)
```

## Introduction

The R ecosystem provides a wide range of options to create different elements in dashboard type 
apps in R-Shiny. Whilst this range of choice can be great for creating anything you want in any 
style you favour, it also leads to a wide range of pitfalls that can undermine your work.

This document outlines a range of packages and elements to avoid and the reasons why, but the key
underlying considerations around many of these are:

- Continuity: 
  - Has the package been recently updated?
  - Are the issues and pull requests raised on the repository getting responded to quickly (if code is visible on GitHub)?
  - Are functions and packages expected to have long term support from their developers?
  - Do functions and packages contain complexities that will be difficult to pick up for whoever 
  succeeds you?
  - If it relies on other packages, is it using the latest version of those
- Accessibility:
  - Are the elements created by functions and packages assigned the appropriate HTML tags and 
  labels?
  - Do elements created by functions and packages work well with keyboard navigation when in 
  screen reader mode?
- Styling
  - Do functions and packages allow you to create elements that are consistent with the [Government Digital Service style guide](https://www.gov.uk/guidance/style-guide).
- Performance / efficiency
  - Do they pull in lots of extra dependencies and code you don't need?
  - How fast are the functions to run?
  
## Packages to avoid

### shinydashboard

[shinydashboard](https://rstudio.github.io/shinydashboard/index.html) is now superseded and has been minimally maintained since 2018. It creates dashboard
elements based on the web toolkit [Bootstrap v3](https://getbootstrap.com/docs/3.3/), which has 
itself moved on to v5. This presents potential issues around continuity, accessibility and 
functionality (in particular when moving to smaller screen sizes). As such, we recommend using 
[bslib](https://rstudio.github.io/bslib/), which provides equivalent functionality, but is underpinned 
by [Bootstrap v5](https://getbootstrap.com/docs/5.3/getting-started/introduction/), which allows 
for better usability across different screen sizes (including high zoom on a computer screen).

### shinyWidgets

[shinyWidgets](https://dreamrs.github.io/shinyWidgets/index.html) provides a wide range of elements for use in Shiny that look good and allow some 
useful advanced / customisable options. However, a large proportion of the widgets provided by 
shinyWidgets do not assign appropriate html tags and labelling. As part of accessibility testing of
DfE dashboards, shinyWidgets elements have consistently been flagged as problematic. They have
confused testers using assistive technology, blocking them from being able to navigate through
dashboards.

Given the above, we recommend avoiding shinyWidgets in DfE Shiny applications.

### Plotly

[Plotly](https://plotly.com/r/) offers a range of functionality to add interactivity to charts. However, these come with the following issues:

- Poor for accessibility
- Defaults add a lot of elements that are mostly distracting / unhelpful
- Tends to be very intensive and if used for lots of charts can really hit performance

We generally advise either keeping chart interactivity to the essentials and using ggiraph in place of Plotly.

## Functions to avoid / use sparingly

### Conditional panels

Whilst [`conditionalPanel()`](https://shiny.posit.co/r/reference/shiny/1.4.0/conditionalpanel.html) 
is useful in some cases, many dashboard creators have had an over-reliance on these elements. 
Conditional panels, when used inappropriately, can lead to unnecessary code complexity and poor 
performance. 

Specifically, the app still processes all the code contained within a conditional panel and then just hides the appropriate elements based on the condition. So if you do 5 different charts in a conditional panel, that's 5 charts your app is building each time, instead of just 1 if built with other methods.

The alternatives to consider first are:

- Dynamically update individual objects such as text, charts and tables within the `render*({})` environments.
- Use `updateCheckboxGroupInput()`, `updateRadioButtons()`, `updateSelectInput()`, etc. functions for 
dynamically updating labels and options in a range of individual inputs.
- `uiOutput()` with `renderUI()` for extended dashboard components, although these should also be 
used with caution.
- Create modules for more complex UIs.

### Select input component with (multiple=TRUE)

The [GOV.UK Design System](https://design-system.service.gov.uk/) explicitly avoids the use of allowing multiple selections via the select input component, due to "[a history of poor usability and assistive technology support](https://design-system.service.gov.uk/components/select/)". One particular set of testing reported the following in reference to the select multiple element.

> Not a single participant relying on the keyboard or a desktop screen reader managed to select multiple non-contiguous options...

As such, the recommendation from the GOV.UK Design System if you need to a user to pick more than one item from a list, is to use an alternative such as a list of checkboxes wherever possible.

Note that the shinyGovstyle package's [`select_Input()`](https://dfe-analytical-services.github.io/shinyGovstyle/reference/select_Input.html) function provides the appropriate digital styling for the select component, whilst also limiting dashboard creators to only single selection inputs.
